---
title: "Data Cleaning"
author: "Alfonso Rojas-Alvarez"
date: "9/26/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### This function will clean up your environment
rm(list = ls())

### Load the packages you will need for your work (you can also do this as you go).
library(dplyr)
library(ggplot2)
library(foreign)

### Set the working directory
setwd("/Users/ajr3347/r_workshop")
```

## Data Cleaning

Let's go over a set of functions that will help you clean data very effectively, and ease the transition from R into Stata. First, let's load our data:

```{r, echo = TRUE}
### For this function, read.delim, we will need to load the foreign package.
wages <- read.delim("urbup.dat", sep = "")

### Give the variables names (some data sets already have them, check first)
names(wages) <- c("wage", "educ", "female", "faed")
```

## Filter

The filter() function allows you to subset rows in your data based on a condition. You can either do this as a separate object, or entirely replace your original data set.

Here, let's first filter female observations:

```{r, echo=TRUE}

### In a new object (notice it carries the variables names, nice!)
female_w <- filter(wages, female == 1)

### Replace the one you had before
wages <- filter(wages, female == 1)

### But that's not always very helpful, since you 
### want the original data set to be available. Let's load it again:
wages <- read.delim("urbup.dat", sep = "")
names(wages) <- c("wage", "educ", "female", "faed")
```

Let's now filter with two conditions, use the & or the | operators for 'and' and 'or' conditions.

```{r, echo=TRUE}

### Let's filter males with low education
males_lowed <- filter(wages, female == 0 & educ < 15)

### Say don't want that object in the environment anymore. 
### To delete an object:
rm(males_lowed)
```

Plot wages:

```{r, echo=TRUE}

### Ask R to treat female variable as a factor 
### Factor: 0 or 1, rather than any decimal number from 0 to 1.
wages$female <- factor(wages$female)
plot1 <- ggplot(data = wages, aes(wage, educ, color=female)) + geom_point() +
  labs(x="Wages", y="Years of Education") + ggtitle("Plot of Wages by Education Level")
plot1
```

## Select

The select() function allows you to subset columns of your data. This is helpful if you want to remove a group of columns when you have a wide array or variables.

```{r, echo=TRUE}

### Select just two variables from the data set
faed_wage <- select(wages, wage, faed)

### You can also remove variables
fem_educ <- select(wages, -wage, -faed)
```

## Mutate

The function mutate() allows you to create a new variable based on a function of an existing one:

```{r}
### Create a new variable that is the log of wages
wages <- mutate(wages, logwage = log(wage))

### Mutate stacks up!
wages <- mutate(wages, logwage_summed = logwage + 123, logwage_summed2 = logwage_summed^2)
```

Let's remove those variables for now:

```{r}
wages$logwage <- NULL
wages$logwage_summed <- NULL
wages$logwage_summed2 <- NULL
```

## Arrange

The function arrange() allows you to 'order' a data set based on a variable, in ascending or descending order.

```{r}

### Ascending order
wages <- arrange(wages, faed)

### Descending order
wages <- arrange(wages, desc(faed))
```

## Summarize

Very simple function to give you aggregates of the data frame, it becomes more powerful when combined with other functions.

```{r}
summarize(wages, mean(wage))
```
## Group By

This function is the one that makes summarize powerful. You can group your data set by a variable, and then obtain summaries based on it:

```{r}
summarize(group_by(wages, faed), mean(wage))
```
The combination of these functions will allow you to do very powerful and elegant cleaning for your data sets. As always, practice makes the master!


