---
title: "Models"
author: "Alfonso Rojas-Alvarez"
date: "9/28/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### This function will clean up your environment
rm(list = ls())

### Load the packages you will need for your work (you can also do this as you go).
library(dplyr)
library(ggplot2)
library(foreign)
library(stargazer)

### Set the working directory
setwd("/Users/Alfonso/Google Drive/UT/Fall 2018/R Workshop")

### For this function, read.delim, we will need to load the foreign package.
wages <- read.delim("urbup.dat", sep = "")

### Give the variables names (some data sets already have them, check first)
names(wages) <- c("wage", "educ", "female", "faed")
```

Note: these models are not necessarily the best or most suited for this particular data set. The data set is used merely for illustrative purposes. Remember there must be a theoretical or methodological justification to use one model or another.

Basic OLS:

```{r}
### Create the regression object
model1 <- lm(wage ~ female + educ + faed, data = wages)

### Summarize to view the results
summary(model1)
```

No constant:

```{r}
model2 <- lm(wage ~ female + educ + faed - 1, data = wages)
summary(model2)
```

Test Normality:

```{r}
### A q-q normal plot compares the results of the 
### ols with a theoretical standard normal distribution
qqnorm(wages$educ)
```

Confidence Interval:

```{r}
confint(model1)
```

GLM:

```{r}
### For a GLM, change the link function 'family'
model3 <- glm(female ~ educ + wage + faed, data=wages, family = binomial(link = "logit"))
summary(model3)
```

Difference in differences:

```{r}
### Difference in Difference is basically just a simple OLs
### What changes is how you code the variables
model4 <- lm(formula = wage ~ female * faed, data = wages)
summary(model4)
```

Fixed Effects:

```{r}
# model5 <- plm(wage ~ educ + faed + female, data = wages, index = c("state", "year"), model = "within", effect= "individual")
# summary(model5)

### I am not running this one since something went wrong with my R version and it is not reading plm
```

Propensity Score Matching:

```{r}
#model6  = matchit(y ~ x1 + x2 + x3, data = data, method = "nearest", ratio = 1)
# summary(model6)

# Same problem as before
```

```{r}
stargazer(model1, model2, model3, model4)

### This will shoot out a chunk of code.
### Copy and paste it into a latex file to get a pretty table!
```


Monte Carlo Simulation:

```{r}
### Generate a distribution
mcsim <- rnorm(50, mean = 0, sd = 1)
```

To Bootstrap a MC Simulation:

```{r}
### Create a function to generate a distribution
rep <- function(i) {
  a <- rbinom(size = 1, n = 720, p = 0.2166)
  mean(a)
}

### Replicate the function 50000 times!
y <- data.frame((sapply(1:50000, rep)))

### Plot it!

a <- ggplot(data=y, aes(y$X.sapply.1.50000..rep..)) + 
  geom_histogram(aes(y =..density..), 
                 fill="red", 
                 alpha = .5) + 
  geom_density(col=2) + 
  labs(x="Whatever you simulated in the X axis", y="Proportion of Replications" )

a
```

Quantiles:

```{r}
q <- c(0.025, 0.975)
quantile <- quantile(mcsim, probs = q)
quantile
```

T-tests:

```{r}
t.test(wages$wage~wages$female) # where y is numeric and x is a binary factor
t.test(wages$wage,wages$educ) # where y1 and y2 are numeric
t.test(wages$wage,wages$educ,paired=TRUE) # where y1 & y2 are numeric
t.test(wages$wage,mu=3) # Ho: mu=3
```

Predict from models:

```{r}
male = c(16, 0, 10)
female = c(16, 1, 10)
newdata <- data.frame(rbind(male, female))
names(newdata)[1]<-paste("educ")
names(newdata)[2]<-paste("female")
names(newdata)[3]<-paste("faed")

### Two specific cases
predict(model1, newdata=newdata)

### Or predict the entire X axis

predicted <- predict(model1)

### To plot them, create a data frame
frame <- data.frame(cbind(predicted, wages$wage))

### Create the plot
plot2 <- ggplot(data=frame, aes(frame$predicted, frame$V2)) + 
  geom_point() + 
  labs(x="Predicted", y="Observed" )
plot2
```

